---
title: "Using proprietary API backends (OpenAI, Anthropic, Gemini, OpenRouter)"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using proprietary API backends (OpenAI, Anthropic, Gemini, OpenRouter)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
```

`screenllm` was designed around a locally-served Ollama ensemble, but
the backend interface is small enough that other providers slot in
cleanly. This vignette shows how to use OpenAI, Anthropic Claude,
Google Gemini, and OpenRouter, either on their own or mixed into an
ensemble alongside local Ollama models.

None of the chunks below run at package-build time (they require
live API keys). Copy them into your R console after setting the
relevant environment variables.

## Store your API keys

Put your keys in `~/.Renviron` so they aren't committed with your
code and aren't visible in `history`. Restart R after editing:

```
OPENAI_API_KEY=sk-...
ANTHROPIC_API_KEY=sk-ant-...
GOOGLE_API_KEY=AIzaSy...
OPENROUTER_API_KEY=sk-or-...
```

Then in R:

```{r}
library(screenllm)

# Each backend reads its key from the corresponding env var by default.
ok  <- backend_openai()
cl  <- backend_anthropic()
gm  <- backend_gemini()
orb <- backend_openrouter(x_title = "screenllm")
```

## Single-provider ensemble

Any backend plugs into `custom_ensemble()`. Model tags follow each
provider's own naming:

```{r}
gpt_ensemble <- custom_ensemble(
  models = c("gpt-5", "gpt-4o"),
  replicates = 1L,
  backend = backend_openai()
)

claude_ensemble <- custom_ensemble(
  models = "claude-sonnet-5",
  replicates = 3L,
  backend = backend_anthropic()
)

gemini_ensemble <- custom_ensemble(
  models = "gemini-2.5-flash",
  replicates = 1L,
  backend = backend_gemini()
)
```

Then `rank_records()` uses the ensemble like any other:

```{r}
records <- read_records("my_corpus.csv")
criteria <- define_criteria(
  scope = "Field-based coral reef restoration",
  inclusions = c("Study is field-based.",
                 "Study describes an intervention.")
)
ranked <- rank_records(records, criteria, ensemble = gpt_ensemble)
```

## OpenRouter: one key, many models

OpenRouter proxies hundreds of models behind an OpenAI-compatible
API, so you can hit a mixed set with a single key and pay one bill.
Model tags are `provider/name`:

```{r}
mixed <- custom_ensemble(
  models = c(
    "openai/gpt-5",
    "anthropic/claude-sonnet-5",
    "google/gemini-2.5-flash",
    "meta-llama/llama-3.1-70b-instruct"
  ),
  replicates = 1L,
  backend = backend_openrouter(x_title = "screenllm")
)
```

The full model catalogue is at <https://openrouter.ai/models>.

## Mixing local Ollama + cloud APIs

The ensemble diversity thesis in Spillias et al. (2026) says that
averaging over *different* models beats averaging replicates of the
*same* model. That works across providers too. A cheap way to add
a cloud model to a mostly-local ensemble is to run two ensembles in
succession over the same corpus and average their per-record scores
manually, or to use OpenRouter as one of two backends:

```{r}
local_ensemble <- default_ensemble_light()   # 4 x Ollama, ~10 GB
cloud_ensemble <- custom_ensemble(
  models = "openai/gpt-5",
  replicates = 1L,
  backend = backend_openrouter()
)

# Two passes; cache_dir shared so each ensemble only scores its own calls.
ranked_local <- rank_records(records, criteria, ensemble = local_ensemble,
                              cache_dir = "cache/")
ranked_cloud <- rank_records(records, criteria, ensemble = cloud_ensemble,
                              cache_dir = "cache/")

# Combine per-record scores.
combined <- merge(
  ranked_local[, c("id", "universal_best_score")],
  ranked_cloud[, c("id", "universal_best_score")],
  by = "id", suffixes = c(".local", ".cloud")
)
combined$universal_best_score <- rowMeans(
  combined[, c("universal_best_score.local", "universal_best_score.cloud")],
  na.rm = TRUE
)
```

A first-class "mixed backend" ensemble is on the roadmap; for now,
this two-pass pattern gives you the same result.

## Cost planning

The paid backends are billed per-token. Rough per-call cost for the
screening prompt (~500 input tokens + ~200 output tokens) at
2026-Q3 prices:

| Model                       | Approx. cost / call |
|-----------------------------|---------------------|
| gpt-5                       | $0.005              |
| gpt-4o-mini                 | $0.0002             |
| claude-sonnet-5             | $0.007              |
| claude-haiku-4-5            | $0.0008             |
| gemini-2.5-flash            | $0.0004             |
| gemini-2.5-pro              | $0.005              |

A 1000-record corpus screened with a 4-model ensemble at 3 replicates
is 12000 calls, so `gpt-5` would cost ~$60; `gemini-2.5-flash` would
cost ~$5. Check each provider's current pricing before running a big
job.

## Troubleshooting

- **`No API key for OpenAI`**: the corresponding env var isn't set.
  Restart R after editing `~/.Renviron`.
- **`HTTP 401`**: bad key. Regenerate from the provider dashboard.
- **`HTTP 429`**: rate-limited. Wait, or set fewer replicates.
  Rate-limit-aware batching is on the roadmap.
- **`relevance field missing or out of range`**: the model returned
  something other than the expected JSON schema. Retry; if it
  persists, try a bigger model.
